---
title: "Quiz 1"
author: "David F. Severski (@dseverski)"
date: "October 18, 2015"
output: html_document
---

```{r, echo=FALSE}
library(dplyr)
```

# Question 1

After untaring the the dataset, how many files are there (including the documentation pdfs)?

Answer: 7

```{r}
dir('./data/yelp_dataset_challenge_academic_dataset') %>% length
```

# Question 2

The data files are in what format?

Answer: JSON

# Question 3

How many lines of text are there in the reviews file (in orders of mangnitude)?

Answer: one millions


```{r}
dim(reviews)
```

# Question 4

Consider line 100 of the reviews file. I've been going to the gran n Eat for almost XXX years'

Answer: 20

```{r}
reviews[100,"text"]
```

# Question 5
What percentage of teh review are five star reviews?

Answer: 37%

```{r}
reviews %>% group_by(stars) %>% 
  tally %>% mutate(pct=(100*n)/sum(n))
```


# Question 6

How many lines are there in the businesses file?

Answer: Around 60 thousand

```{r}
dim(business)
```

# Question 7

Conditional on having an response for the attribute "Wi-Fi", how many businesses are reported for having free wi-fi (rounded to the nearest percentage point)? 

Answer: 40%
```{r}
wifi <- data.frame(wifi=bus_attributes$`Wi-Fi`)
wifi %>% group_by(wifi) %>% 
  filter(!is.na(wifi)) %>% 
  tally %>% mutate(pct = (100* n) / sum(n))
```

# Question 8

About 500 thousand

# Question 9

In the tips file on the 1,000th line, fill in the blank: "Consistently terrible ______"

Answer: service

```{r}
tips[1000,]
```

# Question 10

Answer: Brian

```{r}
user_id <- compliments %>% filter(funny >10000) %>% select(row_number) %>% as.numeric
users[user_id,"name"]
```


# Question 11

Create a 2 by 2 cross tabulation table of when a user has more than 1 fans to if the user has more than 1 compliment of type "funny". Treat missing values as 0 (fans or votes of that type). Pass the 2 by 2 table to fisher.test in R. What is the P-value for the test of independence?

Answer: less than .001

```{r}
compliments$many_fans <- users$fans > 1
table(compliments$many_fans, compliments$funny>1)
fisher.test(table(compliments$many_fans, compliments$funny>1))
```